Learning Discrete Bayesian Networks from Continuous Data

### 2016-07-21 Learning_Discrete_Bayesian_Networks_from_Continuous_Data.tex
Add an experiment on a synthetic dataset

We use the Iris dataset to support this discussion.
In the appendix we show that the Bayesian and MDL methods produce equivalent results under a naive Bayes structure.
However, if a subset of the samples are used then the distribution pattern of data samples becomes less obvious, and the two discretization methods have different outcomes.

Figure \ref{fig:iris_four} shows the relation between the number of samples and the number of \todo{I thought this would be likelihood?} discretization edges on each continuous variable.
Samples are drawn randomly from the dataset \tim{without replacement?} and the result of number of edges are averaged over a thousand runs.
For the variables $C$ and $D$, the correct discretization policy is easier to obtain, and both methods locate two discretization edges within thirty samples, one fifth of the full Iris dataset.
For the variables $A$ and $B$, more data samples are required before settling on the final discretization edges.
Under this case, the proposed Bayesian method is more sensitive than the MDL method and produces discretizations with higher cross-validated likelihood.

\begin{figure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_edges_A}}
		%\caption{A}
		%\label{fig:iris_a}
	\end{subfigure}%
	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_edges_B}}
		%\caption{B}
		%\label{fig:iris_b}
	\end{subfigure}

	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_edges_C}}
		%\caption{C}
		%\label{fig:iris_c}
	\end{subfigure}%
	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_edges_D}}
		%\caption{D}
		%\label{fig:iris_d}
	\end{subfigure}
	\caption{Number of \todo{discretization edges} for each continuous variable in the Iris dataset as the number of samples used in training are varied.}
	\label{fig:iris_four}
\end{figure}

>>>>

To support the argument, a synthetic dataset with $800$ samples and under the network structure in Figure~\ref{fig:exp_discuss} is tested. Data samples are generated by the distribution described in Table~\ref{table:distri_exp}: both $P_1$ and $P_2$ take two discrete values with equal probability. $\cX$ samples numeric numbers given the value of $P_1$ and $P_2$. Note that the symbol $\sim U[a,b]$ in Table~\ref{table:distri_exp} represents an uniform distribution from value $a$ to $b$. The optimal discretization policy on $\cX$ should have six edges $\curly{2.0,3.0,4.0,4.5,5.0,6.0}$, since $\cX$ takes numeric values from $1.0$ to $7.5$ and there are seven intervals corresponding to different joint distribution of $P_1$ and $P_2$.

\begin{table}[ht]
	\centering
	\caption{The distribution of variables in Figure~\ref{fig:exp_discuss} for the proposed experiment. $P_1$ and $P_2$ are discrete variable. $\cX$ is continuous valued.}
	\label{table:distri_exp}
	\begin{tabular}{cl}
		\toprule
		Variable               & Distribution                                   \\ \midrule
		\multirow{2}{*}{$P_1$}&$P(p^{(1)}_1) = 0.5$     \\
		                                       & $P(p^{(2)}_1) = 0.5$     \\ \midrule
		\multirow{2}{*}{$P_2$}&$P(p^{(1)}_2) = 0.5$     \\
		                                       &$P(p^{(2)}_2) = 0.5$     \\ \midrule
		\multirow{4}{*}{$\cX$} & $p(\cx \mid p^{(1)}_1,p^{(1)}_2) \sim U[2,6]$\\
		                                     & $p(\cx \mid p^{(1)}_1,p^{(2)}_2) \sim U[1,3]$ \\
		                                     & $p(\cx \mid p^{(2)}_1,p^{(1)}_2) \sim U[4,5]$ \\
		                                     & $p(\cx \mid p^{(2)}_1,p^{(2)}_2) \sim U[4.5,7.5]$\\
		\bottomrule
	\end{tabular}
\end{table}

Figure~\ref{fig:synthetic} shows how the likelihood of tested data and number of discretization edges on $\cX$ vary with respect to the size of trained data. Trained data are drawn randomly from the synthetic dataset. The remaining ones are tested data. If the size of the trained data is small, it is hard to recognize the underlying pattern of data samples and discretize the continuous variable $\cX$ accordingly. In this case, both likelihood of the trained data and number of discretization edges is small. However, since the Bayesian method is more sensitive to the data distribution, it has better performance of the MDL method in terms of likelihood and lays down more discretization edges. As the size of the trained data grows, both methods improve their accuracy and approach to optimal number of discretization edges. The Bayesian method finds the six discretization edges faster than the MDL method. Finally, as the size of the trained data is large enough to capture the data distribution, both methods can locate the six discretization edges and lead to same accuracy to the tested data. The curves in Figure~\ref{fig:synthetic} demonstrate that the Bayesian method is indeed more sensitive to the dataset than the MDL method. Note that all the results in Figure~\ref{fig:synthetic} are averaged over one hundred times.

%We use the Iris dataset to support this discussion.
%In the appendix we show that the Bayesian and MDL methods produce equivalent results under a naive Bayes structure.
%However, if a subset of the samples are used then the distribution pattern of data samples becomes less obvious, and the two discretization methods have different outcomes.

%Figure \ref{fig:iris_four} shows the relation between the number of samples and the number of \todo{I thought this would be likelihood?} discretization edges on each continuous variable.
%Samples are drawn randomly from the dataset \tim{without replacement?} and the result of number of edges are averaged over a thousand runs.
%For the variables $C$ and $D$, the correct discretization policy is easier to obtain, and both methods locate two discretization edges within thirty samples, one fifth of the full Iris dataset.
%For the variables $A$ and $B$, more data samples are required before settling on the final discretization edges.
%Under this case, the proposed Bayesian method is more sensitive than the MDL method and produces discretizations with higher cross-validated likelihood.

\begin{figure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_demo_like}}
		%\caption{A}
		%\label{fig:iris_a}
	\end{subfigure}%
	\begin{subfigure}{0.45\textwidth}
		\centering
		\scalebox{0.6}{\input{Fig_demo_edge}}
		%\caption{B}
		%\label{fig:iris_b}
	\end{subfigure}

	\caption{The average log-likelihood of the trained data and the average number of discretization edges for $\cX$ as the number of trained data is varied. Trained data are drawn randomly from the synthetic dataset. The remaining ones are tested data. Results are averaged over one hundred times.}
	\label{fig:synthetic}
\end{figure}
